Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2017-12-24 paper-ieee-trans.tex
Update paper-ieee-trans.tex

\begin{abstract}
From the advent of Grid technology as the new paradigm of distributed
computing to the current days of Cloud computing models, the continuous need
of new tools and services to match the scientific community requirements has been
addressed in Europe by a various number of European Commission funded software development
and e-Infrastructures creation, operation and management dedicated projects.
% the creation of software development and e-Infrastructure coordination dedicated projects. 
This work examines the evolution of the software engineering methodologies used in past and present
European Commission-funded projects to illustrate how the research on software
reliability has progressed in Europe over the last 15 years following the
foundation of the first project in 2001. In particular, we describe the team
organizational structure (?), techniques and procedures that have guided the Quality
Assurance definitions in order to deliver quality software. We highlight the challenges
and barriers that have been faced throughout these years and how they were overcomed totally or
partially. To conclude, we discuss possible future directions in software
Quality Assurance methodologies for the upcoming projects.
\end{abstract}

@@

\IEEEPARstart{W}hen e-Infrastructures in Europe were first conceived, new
services to access distributed resources -- data, computing, instrumental --
were initially considered. Major investments were then injected in order to
develop distributed solutions that provided a seamless and transparent access
to those resources, available for the researchers. One of the most known examples is Worldwide
Large Hadron Collider Computing Grid (WLCG) global computing infrastructure, powered by the
Particle Physics scientific community, built up to satisfy the computing needs
of processing, storing and analysing the huge quantity of data coming from the Large Hadron
Collider particle accelerator.

Years of major developments followed, mainly focused on providing new
functionalities, hindering the adoption of software engineering methodologies.
The software delivered was by then less tested, resulting in constant
rollbacks and patches applied to production systems. Thereby, the stability and
reliability of the e-Infrastructures was often compromised, affecting
negatively their exploitation by the scientific communities.

@@

Analyzing the European Commission-funded software projects in the last decade
we have observed a continuous increase of the prominence and robustness of
testing and validation procedures. The first of many reasons is related to the
evolution of the software engineering techniques (e.g. creation of the Agile
Manifesto \cite{agile-manifesto}, rise of DevOps practices \cite{zhu}) together
with the parallel advancements in the ICT field with virtualization and
container technologies. Another one is the emergence of automation and
event-response tools.

>>>>

\begin{abstract}
From the advent of Grid technology as the new paradigm of distributed
computing to the current days of Cloud computing models, the continuous need
of new tools and services to match the scientific community requirements has been
addressed in Europe by a various number of European Commission funded software development
and e-Infrastructures creation, operation and management dedicated projects.
% the creation of software development and e-Infrastructure coordination dedicated projects. 
This work examines the evolution of the software engineering methodologies used in past and present
European Commission-funded projects to illustrate how the research on software
reliability has progressed in Europe over the last 15 years following the
foundation of the first project in 2001. In particular, we describe the team
organizational structure, %(what does it mean?)
 techniques and procedures that have guided the Quality
Assurance definitions in order to deliver quality software. We highlight the challenges
and barriers that have been faced throughout these years and how they were overcomed totally or
partially. To conclude, we discuss possible future directions in software
Quality Assurance methodologies for the upcoming projects.
\end{abstract}

@@

%\IEEEPARstart{W}hen e-Infrastructures in Europe were first conceived, new
%services to access distributed resources -- data, computing, instrumental --
%were initially considered. Major investments were then injected in order to
%develop distributed solutions that provided a seamless and transparent access
%to those resources, available for the researchers. One of the most known examples is Worldwide
%Large Hadron Collider Computing Grid (WLCG) global computing infrastructure, powered by the
%Particle Physics scientific community, built up to satisfy the computing needs
%of processing, storing and analysing the huge quantity of data coming from the Large Hadron
%Collider particle accelerator.

\IEEEPARstart{D}uring the last 15 years the provision of a European e-Infrastructure has been driven by the 
needs of user communities to access large scale computing and data analysis infrastructure to 
support their research needs as part of their pan-European research collaborations. One of the most 
known examples is Worldwide Large Hadron Collider Computing Grid (WLCG) global computing infrastructure, 
powered by the Particle Physics scientific community, built up to satisfy the computing needs
of processing, storing and analysing the huge quantity of data coming from the Large Hadron
Collider particle accelerator. As a consequence a number of important projects have been run in Europe, supported by the European Commission funding of various research programms (Framework Programms 5th, 6th, 7th, Horizon 2020) in order to
develop distributed solutions that provided a seamless and transparent access to those resources, available for the researchers.

Major developments were done throughout the years, mainly focused on providing new
functionalities, hindering the adoption of software engineering methodologies.
The software delivered was by then less tested, resulting in constant
rollbacks and patches applied to production systems. Thereby, the stability and
reliability of the e-Infrastructures was often compromised, affecting
negatively their exploitation by the scientific communities.

@@

Analyzing the European Commission-funded software projects in the last decade
we have observed a continuous increase of the prominence and robustness of
testing and validation procedures. One of many reasons is related to the
evolution of the software engineering techniques (e.g. creation of the Agile
Manifesto \cite{agile-manifesto}, rise of DevOps practices \cite{zhu}) together
with the parallel advancements in the ICT field with virtualization and
container technologies. Another one is the emergence of automation and
event-response tools.
